---
title: "Introduction to AdvancedGgplot"
author: "Christopher Odoom"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to AdvancedGgplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center"
)
```

## Overview

AdvancedGgplot is a modern data visualization package for R, designed as an alternative to ggplot2 with a fully object-oriented approach using R6 classes. The package provides a flexible, composable API for creating complex visualizations.

## Core Concepts

AdvancedGgplot is built around a few key concepts:

1. **AGPlot**: The main container for visualizations
2. **Layers**: Visual elements that represent data (points, lines, etc.)
3. **Aesthetics**: Mappings between data variables and visual properties
4. **Scales**: Control how data values are mapped to visual properties
5. **Facets**: Create multiple panels based on categorical variables
6. **Themes**: Control non-data visual elements like backgrounds and fonts
7. **Annotations**: Add context with text, arrows, etc.

## Installation

You can install the development version of AdvancedGgplot from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("christopherodoom/AdvancedGgplot")
```

## Basic Usage

Let's start by creating a simple scatter plot:

```{r example1, message=FALSE, eval=FALSE}
library(AdvancedGgplot)

# Create sample data
set.seed(123)
data <- data.frame(
  x = 1:20,
  y = sin(1:20 / 2) + rnorm(20, sd = 0.1),
  z = cos(1:20 / 2) + rnorm(20, sd = 0.1),
  group = rep(c("A", "B"), each = 10)
)

# Create a simple scatter plot
plot <- AGPlot$new(data)
plot$add_layer(geom_point(mapping = list(x = "x", y = "y", color = '"blue"')))
plot$plot()
```

## Combining Multiple Layers

One of the strengths of AdvancedGgplot is the ability to combine multiple layers for complex visualizations:

```{r eval=FALSE}
# Combine points and lines
plot <- AGPlot$new(data)
plot$add_layer(geom_line(mapping = list(x = "x", y = "y", color = '"blue"')))
plot$add_layer(geom_point(mapping = list(x = "x", y = "y", color = '"red"')))
plot$plot()
```

## Statistical Transformations

AdvancedGgplot includes statistical transformations like smoothing:

```{r eval=FALSE}
# Add a smoothed line with confidence interval
plot <- AGPlot$new(data)
plot$add_layer(geom_point(mapping = list(x = "x", y = "y", color = '"black"')))
plot$add_layer(geom_smooth(mapping = list(x = "x", y = "y", color = '"blue"')))
plot$plot()
```

## Statistical Visualization

### Boxplots

Boxplots provide a concise summary of a distribution:

```{r eval=FALSE}
# Create data for statistical plots
stat_data <- data.frame(
  category = rep(c("Group A", "Group B", "Group C"), each = 30),
  value = c(
    rnorm(30, mean = 5, sd = 1),
    rnorm(30, mean = 7, sd = 1.5),
    rnorm(30, mean = 4, sd = 0.8)
  )
)

# Create a boxplot
boxplot <- AGPlot$new(stat_data)
boxplot$add_layer(geom_boxplot(mapping = list(
  x = "category", 
  y = "value",
  fill = '"lightblue"',
  color = '"navy"'
)))
boxplot$plot()
```

### Violin Plots

Violin plots show the distribution shape and are useful for comparing distributions between groups:

```{r eval=FALSE}
# Create a violin plot with the same data
violin_plot <- AGPlot$new(stat_data)
violin_plot$add_layer(geom_violin(mapping = list(
  x = "category", 
  y = "value",
  fill = '"lightgreen"',
  color = '"darkgreen"',
  alpha = 0.7
)))
violin_plot$plot()
```

### Combined Statistical Visualization

You can combine boxplots and violin plots for a more comprehensive view:

```{r eval=FALSE}
# Create a plot with both boxplots and violin plots
combined_stat <- AGPlot$new(stat_data)
combined_stat$add_layer(geom_violin(mapping = list(
  x = "category", 
  y = "value",
  fill = '"lightgreen"',
  alpha = 0.5,
  width = 0.9
)))
combined_stat$add_layer(geom_boxplot(mapping = list(
  x = "category", 
  y = "value",
  fill = '"white"',
  alpha = 0.7,
  width = 0.3
)))
combined_stat$plot()
```

## Faceting

Create multi-panel plots with faceting:

```{r eval=FALSE}
# Prepare data in long format for faceting
long_data <- reshape2::melt(
  data[, c("x", "y", "z")], 
  id.vars = "x", 
  variable.name = "series", 
  value.name = "value"
)

# Create a faceted plot
plot <- AGPlot$new(long_data)
plot$add_layer(geom_line(mapping = list(
  x = "x", 
  y = "value", 
  color = '"blue"'
)))
plot$set_facet(facet_wrap(formula = ~ series, ncol = 2))
plot$plot()
```

## Additional Geometry Types

AdvancedGgplot includes a variety of geometry types:

### Histograms

```{r eval=FALSE}
# Create a histogram
plot <- AGPlot$new(data)
plot$add_layer(geom_histogram(mapping = list(
  x = "y", 
  fill = '"lightblue"', 
  color = '"black"'
)))
plot$plot()
```

### Density Plots

```{r eval=FALSE}
# Create a density plot
plot <- AGPlot$new(data)
plot$add_layer(geom_density(mapping = list(
  x = "y", 
  fill = '"lightblue"', 
  color = '"blue"',
  alpha = 0.5
)))
plot$plot()
```

### Bar Charts

```{r eval=FALSE}
# Create data for bar chart
counts <- data.frame(
  category = c("A", "B", "C", "D"),
  value = c(25, 40, 15, 35)
)

# Create a bar chart
plot <- AGPlot$new(counts)
plot$add_layer(geom_bar(mapping = list(
  x = "category", 
  y = "value",
  fill = '"steelblue"'
)))
plot$plot()
```

## Customization with Themes

Customize your plots with themes:

```{r eval=FALSE}
# Apply a dark theme
plot <- AGPlot$new(data)
plot$add_layer(geom_line(mapping = list(x = "x", y = "y", color = '"white"')))
plot$set_theme(theme_dark())
plot$plot()
```

## Annotations

Add context to your plots with annotations:

```{r eval=FALSE}
# Add text annotation
plot <- AGPlot$new(data)
plot$add_layer(geom_line(mapping = list(x = "x", y = "y", color = '"blue"')))
plot$add_annotation(annotation_text(x = 10, y = 0.8, label = '"Peak"'))
plot$plot()
```

## Conclusion

AdvancedGgplot provides a flexible, object-oriented approach to data visualization in R. Its composable API makes it easy to create complex visualizations while maintaining readability and maintainability. 